---
title: "Lab 13"
author: "Adam Hayes, Erin Omyer, Richard Park, Jasmine Sanchez"
date: "4/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
trafficData <- read_csv("TrafficDeathsPer1000.csv")%>%
  rename(Country = "Traffic mortality per 100,000, age adjusted")
populationDensity <- read_csv("PopulationDensity.csv")%>%
  rename(Country = "Population density (per square km)")

trafficDataReal <- trafficData %>%
  select(Country,`2002`,`2004`)
trafficDataReal$`2002` <- as.numeric(trafficDataReal$`2002`)
trafficDataReal$`2004` <- as.numeric(trafficDataReal$`2004`)

populationDensityReal <- populationDensity %>%
  select(Country,`2002`,`2004`)
populationDensityReal$`2002` <- as.numeric(populationDensityReal$`2002`)
populationDensityReal$`2004` <- as.numeric(populationDensityReal$`2004`)


mergedData <- trafficDataReal %>%
  inner_join(populationDensityReal, by = "Country", suffix = c(".trafficDeath",".popDensity"))

mergedDataReal <- na.omit(mergedData)
```

###Overall Question: Do traffic deaths increase with higher population density?

####Importance:

####Our Data:

####Conclusion:




###Adam's Findings

This map_dbl function is calculating the mean of each column in my trafficDataAdam dataset which represents the mean number of traffic deaths per 100,000 people per country in 2002 and 2004.
```{r, include = TRUE}

trafficDataAdam <- na.omit(trafficDataReal) %>%
  select(`2002`,`2004`)
  
as_tibble(trafficDataAdam)%>%
  map_dbl(mean)
```


This function accepts a data frame and a different function, and loops through the data frame, applying the inputted function to each column.  In my case I took my merged data set with each country's traffic deaths per 100,000 ppl in 2002 and 2004, and population density in 2002 and 2004.  I calculated the mean and standard deviation of each column, showing a good distribution of the data in each column.
```{r, include=TRUE}
mergedDataAdam <- mergedDataReal %>%
  select(2:5)

myColSummary <- function(df,myFun){
  myOut <- vector("double", length(df))
  for(i in seq(df)){
    myOut[i] <- myFun(df[[i]])
  }
  return(myOut)
}
myColSummary(mergedDataAdam,mean)
myColSummary(mergedDataAdam,sd)

```

####Permutation
Null: Difference In Means between(Ratio.2004 and Ratio.2002) = Difference In Means between(Random groupings of the 2)

Alternate:Difference In Means between(Ratio.2004 and Ratio.2002) != Difference In Means between(Random groupings of the 2)

Test Statistic: Sample Mean

```{r}
AdamData <- mergedDataReal %>%
  mutate(Ratio.2002 = `2002.popDensity` / `2002.trafficDeath`, 
         Ratio.2004 = `2004.popDensity` / `2004.trafficDeath`)
#higher the ratio, lower the traffic deaths relative to population density

mean2002 <- AdamData %>%
  select(Ratio.2002)%>%
  summarise(mean = mean(Ratio.2002))
mean2004 <- AdamData %>%
  select(Ratio.2004)%>%
  summarise(mean = mean(Ratio.2004))

differenceActual = mean2004 - mean2002

testData <- AdamData %>%
  select(Ratio.2002, Ratio.2004)

combined <- c(testData$Ratio.2002, testData$Ratio.2004)

permutationAdam <- function(perms = 1000, values, n1)
{
  outputMean <- vector("double", perms)
  for (i in c(1:perms))
  {
    # Randomly separate vector "values" into disjoint 
    # groups of size "n1" and "length(values) - n1" respectively
    
    vector1 <- sample(values, n1)
    vector2 <- setdiff(values, vector1)

    # Compute the sample means for the two groups from 
    firstMean <- mean(vector1)
    secondMean <- mean(vector2)
    
    # Compute the difference in sample means
    meanDifference <- firstMean - secondMean
    
    outputMean[i] <- meanDifference
  }
  # Return new updated vector, created in step 1
  return(outputMean)
}

resultAdam <- permutationAdam(1000,combined,150)

ggplot()+
  geom_histogram(aes(x=resultAdam), col = "red")+
  geom_vline(xintercept = 1.33, col = "blue")+
  labs(title="Distribution of Difference in Means of random groupings of Ratio.2004 and Ratio.2002")+
  xlab("Difference")

percentileData <- sum(resultAdam<1.33)/1000
percentileData
```

####Conclusion
My actual difference in means of the two groups was 1.33, represented by the blue vertical line in the histogram.  As a percentile, it is represented by the decimal number below the histogram. Looking at the graph influences me to accept the alternate hypothesis because it seems as if the distribution of mean differences is so spread out, it is not easily concluded that the vertical line is the center of the data. However, the percentile shows that the actual difference in means of the two groups is very close to the 50th percentile of the permutation data, so therefore I would lean towards accepting the null hypothesis.


###Richard's Findings

The map_dbl function is used here to calculate the mean of each column in my TrafficRichard dataset which represents the median number of traffic deaths per 100,000 people per country in 2002 and 2004.

```{r, include = TRUE}

TrafficRichard <- na.omit(trafficDataReal) %>%
  select(`2002`,`2004`)

as_tibble(TrafficRichard)%>%
  map_dbl(median)
```

The function here accepts a data frame and a different function. From there, it loops through the data frame, applying the inputted function to each column.  In this case, I took my merged data set with each country's traffic deaths per 100,000 ppl in 2002 and 2004, and population density in 2002 and 2004.  I calculated the median and standard deviation of each column, showing a good distribution of the data in each column.

```{r, include=TRUE}
mergedRichardData <- mergedDataReal %>%
  select(2:5)

myColSummary <- function(df,myFun){
  myOut <- vector("double", length(df))
  for(i in seq(df)){
    myOut[i] <- myFun(df[[i]])
  }
  return(myOut)
}
myColSummary(mergedRichardData,median)
myColSummary(mergedRichardData,sd)

```

####Permutation
Null: Difference in Medians between(Ratio.2004 and Ratio.2002) = Difference in Medians between(Random groupings of the 2)

Alternate:Difference in Medians between(Ratio.2004 and Ratio.2002) != Difference in Medians between(Random groupings of the 2)

Test Statistic: Sample Median